\begin{example}
	
	The secretary problem can be framed as a game between a Robot and Nature. Nature presents a sequence of candidates with varying qualities, and the Robot aims to select the best candidate by making sequential decisions.
	
	
	Nature presents $n$ candidates in random order. Let $X_i$ represent the quality of the $i$-th candidate. The qualities $\{X_1, X_2, \ldots, X_n\}$ are assumed to be independently and identically distributed (i.i.d.) random variables.
	
	
	
	The Robot must decide whether to accept or reject each candidate as they are presented. The decision rule $U$ determines the stopping point based on the observed qualities.
	
	
	
	The Robot aims to minimize the regret associated with not selecting the best candidate. The cost function $C$ can be defined as follows:
	\begin{equation}
		C(U(\tilde{D}), s) = 1-\delta_{U(\tilde{D}),s},
	\end{equation}
	where $X_{U}$ is the quality of the candidate selected by the decision rule $U$.
	
	
	
	A decision rule $U$ is a function that maps the sequence of observed qualities and the number of remaining candidates to a decision:
	\begin{equation}
		U: \{X_1, X_2, \ldots, X_k\} \times \{k\} \rightarrow \{ \text{Accept}, \text{Reject} \}.
	\end{equation}
	
	
	
	The conditional expected cost given the observed sequence up to candidate $k$ and the decision rule $U$ is:
	\begin{equation}
		\mathbb{E}[C|\tilde{D}, I] = \int ds_{k+1} \cdots ds_n \, C(U, s) p(\{s_{k+1},\dots s_{n}\}|\tilde{D}, I).
	\end{equation}
	
	
	
	The optimal decision rule $U^*$ minimizes the maximum expected cost:
	\begin{equation}
		U^* = \inf_U \sup_{\{X_1, \ldots, X_n\}} \mathbb{E}[C(U, X)|\{X_1, X_2, \ldots, X_n\}, I].
	\end{equation}
	
	
	In the classical secretary problem, the optimal strategy is to observe the first $m = \left\lfloor \frac{n}{e} \right\rfloor$ candidates and then select the first candidate thereafter that is better than all previous candidates.
	
	
	
	The success probability of the classical strategy is approximately $\frac{1}{e}$, which is the probability that the best candidate is selected using this strategy.
	
\end{example}